---
title: "Stats 511 HW 5 - Comparisons among several samples"
author: "Nate Lant"
date: "October 14, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=F}
library(tidyverse)
library(pander)
library(ploty)
```

## Problem 4

## Problem 10


## Problem 16
Obtain the analysis of variance table using the Spock data.
```{r 16 work, include=FALSE}
spock <- read_csv("data/case0502.csv")

aov(Percent ~ Judge, data=spock) %>%
  pander()

```

## Problem 17

## Problem 18
```{r 18setup, include = F}
rat <- read_csv("data/ex0518.csv")
```

(a) Obtain estimated means for the model with six independent samples, one for each treatment. Determine the residuals and plot them versus the estimated means. Plot the residuals versus the day on which the investigation was conducted. Is there any indication that the methods of this chapter are not appropriate? 

**Answer:** The estimated means are reported below. The residuals, plottted versus the estimated means, are also displayed below. 
```{r setup18a, echo=F}
# Estimated means
rat %>% select(Protein, Treatment) %>%
  group_by(Treatment) %>%
  summarise(mean = mean(Protein)) %>%
  pander()

# Estimated means in a vector
rat.means <- rat %>% select(Protein, Treatment) %>%
  group_by(Treatment) %>%
  summarise(mean = mean(Protein)) %>% c()

# Residuals
rat.resid <- lm(Protein ~ Treatment, data = rat) %>%
  resid()

ggplot(rat, aes(x=rat.means$mean, y=))



# Boxplot visualization
rat %>% ggplot(aes(x=Treatment, y=Protein, shape=Day, color=TrtDayGroup)) +
  geom_boxplot(alpha=0.1) + geom_jitter(width = 0.1, size=4) %>%
  ggploty()


```

(b) Obtain estimated means for the model with 10 independent samples, one from each treatment-day combination. Calculate the ANOVA F-test to see whether these 10 groups have equal means. 

(c) Use (a) and (b) and the methods of Section 5.4.1, to test whether the means for the control groups on different days are different. That is, compare the model with 10 different means to the model in which there are 6 different means.


## Problem 23





