---
title: "HW 7"
author: "Nate Lant"
date: "October 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("car")
```

## Problem 2
Big Bang Data. Explain why improved measurement of distance would lead to more precise estimates of the regression coefﬁcients.

**Answer** It is hard to measure a relationship between distance and velocity if the measurement of distance is inaccurate. The Y variable (distance) is described as a function of the explanitory variable X (velocity). While velocity is accurately measured, it cannot explain the response variable without accurate measurements of the distance.

## Problem 9


## Problem 15
(a) Draw a scatterplot of proportion of pollen removed versus duration of visit, for thebumblebee queens. 

```{r a bees, include=FALSE}
types <- list('n', 'n', 'c')
bees <- read_csv("data/ex0327.csv",
                 col_names = c("Pollen", "Hours", "Bee"), 
                 skip = 1)

# Use this code...
queenbees <- filter(bees, Bee == "Queen")
```

(b) Fit the simple linear regression of proportion of pollen removed on duration of visit. Draw the estimated regression line on the scatterplot.

```{r b bees, echo=FALSE}
(bees.lm <- lm(queenbees$Pollen ~ queenbees$Hours))
plot(queenbees$Hours, queenbees$Pollen)
abline(bees.lm)
```

## Problem 21
```{r planets setup, include=FALSE}
planet <- read_csv("data/ex0721.csv")
(planet.lm <- lm(planet$Distance ~ planet$Order))
```
### First, consider the planets without the astroid belt.
(a) draw a scatterplot of log of distance versus order
(b) include the least squares estimated simple linear regression line on the plot
```{r planets a-c, echo=FALSE}
Distance <- log(planet$Distance)
plot(planet$Order, Distance)
(planet.lm <- lm(Distance ~ planet$Order))
abline(planet.lm)

```

(c) ﬁnd the estimate of standard deviation from the least squares ﬁt

***Answer*** the residual standard error is `r round(summary(planet.lm)$sigma, 3)`.

(d) draw a scatterplot of the residuals versus the ﬁtted values from this ﬁt. 
```{r planet d, echo=FALSE}
plot(fitted.values(planet.lm),planet.lm$residuals)
```

### Now consider the data including the astroid belt.
(e) draw a scatterplot of log of distance versus order, 
(f) include the least squares estimated simple linear regression line on the plot, 
```{r planets e-f, echo=FALSE}
Distance2 <- log(planet$Distance2)
plot(planet$Order2, Distance2)
(planet.lm2 <- lm(Distance2 ~ planet$Order2))
abline(planet.lm2)
```

(g) ﬁnd the estimate of sigma from the least squares ﬁt

***Answer*** the residual standard error is `r round(summary(planet.lm2)$sigma, 3)`.

(h) draw a scatter plot of the residuals versus the ﬁtted values from this ﬁt. 
```{r planet h, echo=FALSE}
plot(fitted.values(planet.lm2),planet.lm2$residuals)
```

(i) Does it appear that the simple linear (straight line) regression model ﬁts better to the ﬁrst set of 9 planets or the second set of 10 “planets”? Explain. 


## Problem 25

## Problem 29



