---
title: "HW 8"
author: "Nate Lant"
date: "10/28/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 6


## Problem 12


## Problem 19
```{r 6setup, include=FALSE}
pollen <- read_csv("data/ex0327.csv")
queen <- pollen %>%
  filter(BeeType == "Queen")
```
Reconsider the pollen removal data of Exercise 3.27 and the regression of pollen removed on time spent on ﬂower, for the bumblebee queens only. 

***Answer:***
(a) What problems are evident in the residual plot? 
```{r bees 6a, echo=FALSE}
pollenremoved <- queen$PollenRemoved
timeonflower <- queen$DurationOfVisit

# Plot the residuals
bees.residual <- resid(lm(pollenremoved ~ timeonflower, data = queen))
plot(bees.residual ~ timeonflower)
```
The problem with the residuals plot is that there are major outliers.

(b) Do log transformations of Y or X help any? 
```{r bees 6b, echo = FALSE}
pollenremoved.log <- log(queen$PollenRemoved)
timeonflower.log <- log(queen$DurationOfVisit)

# Plot various transformation residual plots
## Plot log of explanitory variable = time on flower
bees.residual.logx <- resid(lm(pollenremoved ~ timeonflower.log, data = queen))
plot(bees.residual.logx ~ timeonflower.log)

# Plot log of the response variable
bees.residual.logy <- resid(lm(pollenremoved.log ~ timeonflower, data = queen))
plot(bees.residual.logy ~ timeonflower)

```

(c) Try ﬁtting the regression only for those times less than 31 seconds (i.e., excluding the two longest times). Does this ﬁt better? 
```{r bees 6c, echo=FALSE}
queen.edited <- queen %>% 
  filter(!(queen$DurationOfVisit >= 31))

timeonflower.edited <- queen.edited$DurationOfVisit
pollenremoved.edited <- queen.edited$PollenRemoved

plot(pollenremoved.edited ~ timeonflower.edited)
abline(lm(pollenremoved.edited ~ timeonflower.edited))
```

## Problem 20


## Problem 25